---
title: "Exercise 4"
output:
  pdf_document: default
  html_notebook: default
author: Mads Nielsen & Daniel Winkler
---
First we write a loop that generates the necessary values for a. and c. For the comparison of the quantiles we repeat the calculation 10 times in order to gain accuracy and still keep computation time reasonable. 

```{r eval=FALSE}
N <- 10000
set.seed(1234)
reps <- 10
ntot <- reps * N

Ts <- c(25, 50, 100, 250, 500)
t25_store <- matrix(NA, Ts[1], 1)
t50_store <- matrix(NA, Ts[2], 1)
t100_store <- matrix(NA, Ts[3], 1)
t250_store <- matrix(NA, Ts[4], 1)
t500_store <- matrix(NA, Ts[5], 1)
q25_store <- matrix(NA, reps, 2)
q50_store <- matrix(NA, reps, 2)
q100_store <- matrix(NA, reps, 2)
q250_store <- matrix(NA, reps, 2)
q500_store <- matrix(NA, reps, 2)

for(Ti in Ts){
  t_store <- matrix(NA, N, 1)
  quantile_store <- matrix(NA, reps, 2)
  for(qrep in 1:reps){
    for(nrep in 1:N){
      # Draw randomness
      eps <- rnorm(Ti)
      # Generate Data
      Y <- matrix(NA, Ti, 1)
      Y[1,1] <- eps[1]
      for (irep in 2:Ti){
        Y[irep, 1] <- Y[irep-1,1] + eps[irep] 
      }
      # Generate first difference
      dY <- matrix(NA, (Ti-1), 1)
      for (jrep in 1:(Ti-1)){
        dY[jrep, 1] <- Y[jrep + 1, 1] - Y[jrep, 1]
      }
      # Adjust data length for regression
      Yreg <- Y[1:nrow(dY), 1]
      # Do the regression to obtain t-test 
      mod <- lm(dY~0+Yreg)
      t_store[nrep, 1] <- summary(mod)$coefficients[3]
    }
    quantile_store[qrep, ] <- quantile(t_store, probs = c(0.01, 0.05))
  }
  if(Ti == 25){
    t25_store <- t_store
    q25_store <- quantile_store
  }else if(Ti == 50){
    t50_store <- t_store
    q50_store <- quantile_store
  }else if(Ti == 100){
    t100_store <- t_store
    q100_store <- quantile_store
  }else if(Ti == 250){
    t250_store <- t_store
    q250_store <- quantile_store
  }else if(Ti == 500){
    t500_store <- t_store
    q500_store <- quantile_store
  }
  print(Ti)
}
```

The valus seem reasonably close to those in Maddala and Kim. 

```{r}
library(knitr)
load("cvals.Rdata")
row.names(cvals) <- c("25", "50", "100", "250", "500")
kable(cvals, col.names = c("1%", "5%"))
```

For c.1) and c.2) only 5 repetitions are considered because of computation time. Furthermore, N is reduced to 5000.

```{r, eval=FALSE}
N <- 5000
set.seed(1234)
reps <- 5
ntot <- reps * N

Ts <- c(25, 50, 100, 250, 500)
t25_store <- matrix(NA, Ts[1], 1)
t50_store <- matrix(NA, Ts[2], 1)
t100_store <- matrix(NA, Ts[3], 1)
t250_store <- matrix(NA, Ts[4], 1)
t500_store <- matrix(NA, Ts[5], 1)
q25_store <- matrix(NA, reps, 2)
q50_store <- matrix(NA, reps, 2)
q100_store <- matrix(NA, reps, 2)
q250_store <- matrix(NA, reps, 2)
q500_store <- matrix(NA, reps, 2)

for(Ti in Ts){
  t_store <- matrix(NA, N, 1)
  quantile_store <- matrix(NA, reps, 2)
  for(qrep in 1:reps){
    for(nrep in 1:N){
      # Draw randomness
      eps <- rnorm(Ti)
      # Generate Data
      Y <- matrix(NA, Ti, 1)
      Y[1,1] <- eps[1]
      for (irep in 2:Ti){
        Y[irep, 1] <- 5 + Y[irep-1,1] + eps[irep] 
      }
      # Generate first difference
      dY <- matrix(NA, (Ti-1), 1)
      for (jrep in 1:(Ti-1)){
        dY[jrep, 1] <- Y[jrep + 1, 1] - Y[jrep, 1]
      }
      # Adjust data length for regression
      Yreg <- Y[1:nrow(dY), 1]
      # Do the regression to obtain t-test 
      mod <- lm(dY~Yreg)
      t_store[nrep, 1] <- summary(mod)$coefficients[2,3]
    }
    quantile_store[qrep, ] <- quantile(t_store, probs = c(0.01, 0.05))
  }
  if(Ti == 25){
    t25_store <- t_store
    q25_store <- quantile_store
  }else if(Ti == 50){
    t50_store <- t_store
    q50_store <- quantile_store
  }else if(Ti == 100){
    t100_store <- t_store
    q100_store <- quantile_store
  }else if(Ti == 250){
    t250_store <- t_store
    q250_store <- quantile_store
  }else if(Ti == 500){
    t500_store <- t_store
    q500_store <- quantile_store
  }
  print(Ti)
}
```


```{r}
# cvals <- rbind(colMeans(q25_store), colMeans(q50_store), colMeans(q100_store), colMeans(q250_store), colMeans(q500_store))
# save(cvals, file = "cvalsConst.Rdata")
load("cvalsConst.Rdata")
row.names(cvals) <- c("25", "50", "100", "250", "500")
kable(cvals, col.names = c("1%", "5%"))
```




```{r, eval = FALSE}
N <- 5000
set.seed(1234)
reps <- 5
ntot <- reps * N

Ts <- c(25, 50, 100, 250, 500)
t25_store <- matrix(NA, Ts[1], 1)
t50_store <- matrix(NA, Ts[2], 1)
t100_store <- matrix(NA, Ts[3], 1)
t250_store <- matrix(NA, Ts[4], 1)
t500_store <- matrix(NA, Ts[5], 1)
q25_store <- matrix(NA, reps, 2)
q50_store <- matrix(NA, reps, 2)
q100_store <- matrix(NA, reps, 2)
q250_store <- matrix(NA, reps, 2)
q500_store <- matrix(NA, reps, 2)

for(Ti in Ts){
  t_store <- matrix(NA, N, 1)
  quantile_store <- matrix(NA, reps, 2)
  for(qrep in 1:reps){
    for(nrep in 1:N){
      # Draw randomness
      eps <- rnorm(Ti)
      # Generate Data
      trend <- 1:Ti
      Y <- matrix(NA, Ti, 1)
      Y[1,1] <- eps[1] + trend[1]
      for (irep in 2:Ti){
        Y[irep, 1] <- 5 + trend[irep] + Y[irep-1,1] + eps[irep] 
      }
      # Generate first difference
      dY <- matrix(NA, (Ti-1), 1)
      for (jrep in 1:(Ti-1)){
        dY[jrep, 1] <- Y[jrep + 1, 1] - Y[jrep, 1]
      }
      # Adjust data length for regression
      Yreg <- Y[1:nrow(dY), 1]
      trendreg <- trend[1:nrow(dY)]
      # Do the regression to obtain t-test 
      mod <- lm(dY~Yreg+trendreg)
      t_store[nrep, 1] <- summary(mod)$coefficients[2,3]
    }
    quantile_store[qrep, ] <- quantile(t_store, probs = c(0.01, 0.05))
  }
  if(Ti == 25){
    t25_store <- t_store
    q25_store <- quantile_store
  }else if(Ti == 50){
    t50_store <- t_store
    q50_store <- quantile_store
  }else if(Ti == 100){
    t100_store <- t_store
    q100_store <- quantile_store
  }else if(Ti == 250){
    t250_store <- t_store
    q250_store <- quantile_store
  }else if(Ti == 500){
    t500_store <- t_store
    q500_store <- quantile_store
  }
  print(Ti)
}
```

```{r}
#cvals <- rbind(colMeans(q25_store), colMeans(q50_store), colMeans(q100_store), colMeans(q250_store), colMeans(q500_store))
#save(cvals, file = "cvalsTrend.Rdata")
load("cvalsTrend.Rdata")
row.names(cvals) <- c("25", "50", "100", "250", "500")
kable(cvals, col.names = c("1%", "5%"))

```

