---
title: "Problem 5"
output:
  pdf_document: default
  html_notebook: default
author: Mads Nielsen & Daniel Winkler
---
```{r,include=FALSE} 
#SETUP FOR CODE
rm(list=ls())
library(readxl)
library(forecast)
library(tseries)
library(TSA)
setwd("/home/daniel/pCloudDrive/Aarhus/TS/Data")
gdp<-read_excel("US_GDP.xlsx")
```
Looking at the PACF and the ACF
```{r}
lgdp<-ts(gdp$`Log US_GDP`, start=1947, frequency = 4)
acf(lgdp, lag.max=262)
pacf(lgdp)
```
The ACF plot shows sign of some business cycle correlation. We want to check if the process is stationary
```{r}
adf.test(lgdp)
```
So the augmented Dickey Fuller test yields a high p-value, and thus we cannot reject the null that there is a unit root. 
This is also evident from exploring the plot of the series.
```{r}
ts.plot(lgdp)
```

This certainly does not looks stable. We try differencing the process
```{r}
difflgdp<-diff(lgdp)
ts.plot(difflgdp)
acf(difflgdp)
adf.test(difflgdp)
```
We reject the Dickey Fuller of unit root with a very low p-value. The plot does look much more stationary as well. 

We would like to model the process to see if the non-stationarity is caused by seasonality and/or trend. 

```{r}
#making a model
trend <- 1:nrow(gdp)
season <- seasonaldummy(lgdp)
mod<-lm(lgdp~trend+season)
summary(mod)
# residuals are now demeaned, detrended, deseasoned data
tryresid<-resid(mod)
ts.plot(tryresid)
acf(tryresid)
pacf(tryresid)
```

The trend does explain some of the process, but the seasonality isn't very good at explaining the models. 
```{r}
adf.test(resid(mod))
pp.test(resid(mod))
```

And the residuals still are a unit root process.
We could also check the the different periods to see if the estimate is robust
```{r}
y1<-window(lgdp,1947,1957)
y2<-window(lgdp,1957,1967)
y3<-window(lgdp,1967,1977)
y4<-window(lgdp,1977,1987)
y5<-window(lgdp,1987,1997)
y6<-window(lgdp,1997,2007)
y7<-window(lgdp,2007,2012)

ts.plot(y1)
ts.plot(y2)
ts.plot(y3)
ts.plot(y4)
ts.plot(y5)
ts.plot(y6)
ts.plot(y7)


adf.test(y1)
adf.test(y2)
adf.test(y3)
adf.test(y4)
adf.test(y5)
adf.test(y6)
adf.test(y7)
```
It does seem like the unit root is present in all smaller intervals as well. If the process is detrended.
```{r}

#Detrend, demeaned different periods
#Model 1947-1957
trend1 <- 1:length(y1)
mod1<-lm(y1~trend1)
summary(mod1)
ts.plot(resid(mod1))
adf.test(resid(mod1))
pp.test(resid(mod1))

#1957-1967
trend2 <- 1:length(y2)
mod2<-lm(y2~trend2)
summary(mod2)
ts.plot(resid(mod2))
adf.test(resid(mod2))
pp.test(resid(mod2))

#1967-1977
trend3 <- 1:length(y3)
mod3<-lm(y3~trend3)
summary(mod3)
ts.plot(resid(mod3))
adf.test(resid(mod3))
pp.test(resid(mod3))

#1977-1987
trend4 <- 1:length(y4)
mod4<-lm(y4~trend4)
summary(mod4)
ts.plot(resid(mod4))
adf.test(resid(mod4))
pp.test(resid(mod4))

#1987-1997
trend5 <- 1:length(y5)
mod5<-lm(y5~trend5)
summary(mod5)
ts.plot(resid(mod5))
adf.test(resid(mod5))
pp.test(resid(mod5))

#1997-2007
trend6 <- 1:length(y6)
mod6<-lm(y6~trend6)
summary(mod6)
ts.plot(resid(mod6))
adf.test(resid(mod6))
pp.test(resid(mod6))

#2007-2012
trend7 <- 1:length(y7)
mod7<-lm(y7~trend7)
summary(mod7)
ts.plot(resid(mod7))
adf.test(resid(mod7))
pp.test(resid(mod7))
```


